The Castro AMR Simulation Code: Current and Future Developments

### 2019-10-01 astronum_2019_Castro/astronum-castro-2019.tex
some more work

The original approach to parallelism in \castro\ was MPI + OpenMP,
with scaling on manycore architectures achieve using a tiling approach
to OpenMP~\cite{tiling}.  More recently we've ported \castro\ to GPUs
using CUDA, using the same computational kernels as the MPI + OpenMP
version.  This reuse keeps the code base managable---we don't need
separate kernels for each architecture---while allowing us to take
advantage of current and next generation supercomputers.  All of the
solvers needed to run our core science problems run on GPUs: the main
unsplit PPM hydrodynamics scheme~\cite{ppm,millercolella:2002}, self-gravity via multigrid with
isolated boundary conditions, thermal diffusion, and nuclear
reactions.  Our approach is to put the data on GPUs and then run all
of the kernels on the GPUs, minimizing data movement.  This has given
us enormous speed-ups.

\begin{figure}[t]
\centering
\includegraphics[width=0.8\linewidth]{wdmerger_gpu}
\caption{\label{fig:gpu} blah}
\end{figure}

>>>>

The original approach to parallelism in \castro\ was MPI + OpenMP,
with scaling on manycore architectures achieve using a tiling approach
to OpenMP~\cite{tiling,astronum:2017}.  The general approach is to
have MPI distribute the AMR grids across nodes and have OpenMP threads
work on regions of each grid (tiles) by passing the box describing the
tile into the computational kernel.  More recently we've ported
\castro\ to GPUs using CUDA, using the same computational kernels as
the MPI + OpenMP version.  When run on GPUs, each CUDA thread handles
the update of a single zone, simply by passing that zone index into
the computational kernel.  This reuse keeps the code base
managable---we don't need separate kernels for each
architecture---while allowing us to take advantage of current and next
generation supercomputers.  All of the solvers needed to run our core
science problems run on GPUs: the main unsplit PPM hydrodynamics
scheme~\cite{ppm,millercolella:2002}, self-gravity via multigrid with
isolated boundary conditions, thermal diffusion, and nuclear
reactions.  Our approach is to put the data on GPUs and then run all
of the kernels on the GPUs, minimizing data movement.  This has given
us enormous speed-ups.  Figure~ref{fig:gpu} shows weak scaling on the
OLCF Summit machine for a WD merger problem~\cite{wdmergerI}.  When
using CPUs, we use 42 IBM Power 9 cores per node, while when using
GPUs we use 6 Nvidia Volta GPUs per node.  We see that the code
performance is about $10\times$ higher per node when using GPUs, while
both show excellent weak scaling.

\begin{figure}[t]
\centering
\includegraphics[width=0.8\linewidth]{wdmerger_gpu}
\caption{\label{fig:gpu} \castro\ weak scaling on the OLCF Summit
  machine for the white dwarf merger problem.  On each node we use
  either 42 IBM Power 9 CPU cores or 6 Nvidia Volta GPUs.  Ideal
  scaling is shown for both cases as the solid line.}
\end{figure}
